---
title: "Chocolate Chip Cookies"
execute:
  error: true
author: "Jia Yee Wong"
output: html_document
---

## Reading In the Data

First, read in the CSV data of cookie ingredients.
Make sure that your end-result data has appropriate types for each column - these should match the types provided in the documentation in the README.md file.

```{r}
cookie <- read.csv("choc_chip_cookie_ingredients.csv")
head(cookie)

```

```{python}
import pandas as pd
cookie_python = pd.read_csv('choc_chip_cookie_ingredients.csv')
print(cookie_python)
```



## Exploratory Data Analysis

Exploratory data analysis is the process of getting familiar with your dataset. To get started, [this blog post](https://www.mrdbourke.com/a-gentle-introduction-to-exploratory-data-analysis/) provides a nice checklist to get you thinking:

> 1.  What question(s) are you trying to solve (or prove wrong)?
> 2.  What kind of data do you have and how do you treat different types?
> 3.  What's missing from the data and how do you deal with it?
> 4.  Where are the outliers and why should you care about them?
> 5.  How can you add, change or remove features to get more out of your data?

### Generating Questions

Generate at least 5 questions you might explore using this database of cookie ingredients.

1. Which ingredients appear most frequently across cookie recipes?
2. For a single ingredient are the units consistent or does it requires standardization before analysis?
3. Are there any extreme outliers that might indicate data entry errors?
4. Are there any missing data?
5. How does the amount of sugar change as the amount of flour increases, is the relationship linear, or does it vary widely?


### Skimming the Data

One thing we often want to do during EDA is to examine the quality of the data - are there missing values? What quirks might exist in the dataset?

The `skimr` package in R, and the similar `skimpy` package in python (which has a much better name, in my opinion), can help provide visual summaries of the data. 

Install both packages, and read the package documentation ([R](https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html), [Python](https://pypi.org/project/skimpy/)).

[Part 1] Use each package and generate summaries of your data that require the use of at least some non-default options in each package's `skim` function.


```{r}
library(skimr)
skim(cookie)
```

```{python}
from skimpy import skim
skim(cookie_python)
```



[Part 2] Write 1-2 sentences about what you can tell from each summary display you generate. Did you discover anything new about the data?

Both summaries confirm that the dataset has 1990 rows, but both skimpy and skimr show that there are 1010 missing values in the Rating column and a maximum Quantity of 48, suggesting potential outliers.


### Generating Tables

Another useful technique for exploratory data analysis is to generate summary tables. 
You may want to use the `dplyr` package in R (`group_by` or `count` functions), as well as the `groupby` and `count` methods in Pandas. [Python example](https://sparkbyexamples.com/pandas/pandas-groupby-count-examples/), [R example](https://dplyr.tidyverse.org/reference/count.html)

[Part 1] Using R and Python, generate a table that shows what **proportion** of recipes contain each type of ingredient, for the most common 20 ingredients.
```{r}
library(dplyr)

total <- n_distinct(cookie$Recipe_Index)

top_ingredients <- cookie |>
  group_by(Ingredient) |>
  summarise(count = n_distinct(Recipe_Index)) |>
  mutate(proportion = count / total) |>
  slice_max(order_by = proportion, n = 20) 

print(top_ingredients)
```

```{python}
total_recipes = cookie_python['Recipe_Index'].nunique()
ingredient_counts = cookie_python.groupby('Ingredient')['Recipe_Index'].nunique()
proportions = (ingredient_counts / total_recipes).sort_values(ascending=False)

threshold = proportions.iloc[19]

top_ingredients = proportions[proportions >= threshold]

print(top_ingredients)
```


[Part 2] Print out a character string that lists all of the ingredients that do not appear in at least 20 recipes.
```{r}
rare <- cookie |>
  group_by(Ingredient) |>
  summarise(n = n_distinct(Recipe_Index)) |>
  filter(n < 20) |>
  pull(Ingredient)

toString(rare)
```

```{python}
counts = cookie_python.groupby('Ingredient')['Recipe_Index'].nunique()
rare = counts[counts < 20].index.tolist()
print(", ".join(rare))
```



### Visualization

Using whatever plotting system you are comfortable with in R or python, see if you can create a couple of useful exploratory data visualizations which address one of the questions you wrote above - or another question which you've come up with as you've worked on this assignment.

[Part 1] Create at least one plot (it doesn't have to be pretty) that showcases an interesting facet of the data.
```{r}
library(ggplot2)
library(dplyr)

top_20 <- cookie |>
  count(Ingredient, sort = TRUE) |>
  slice_max(order_by = n, n = 20)

ggplot(top_20, aes(x = reorder(Ingredient, n), y = n)) +
  geom_col() +
  labs(
    title = "Top 20 Most Common Cookie Ingredients",
    x = "Ingredient",
    y = "Number of Recipes"
  ) +
  coord_flip()
```



[Part 2] Write 2-3 sentences about what you can learn from that plot and what directions you might want to investigate from here.

The plot shows that eggs, vanilla, and flour are the top 3 ingredients. Looking at the top 20 ingredients also reveals less common additions like walnuts and oats. Based on this, I would like to compare recipes to see how the amounts of the main ingredients affect the final cookie.